# Reproducible Research: Peer Assignment 1
---
```{r setoptions, echo=FALSE, results='hide'}
opts_chunk$set(echo=TRUE, results='asis')
```
## Loads and Preprocess data
Reads the data.
```{r}
if(!file.exists('activity.csv')){
unzip('activity.zip')
}
activity <- read.csv('activity.csv')
```
Creates a date.time column that combines date and interval columns.
```{r}
time <- formatC(activity$interval / 100, 2, format='f')
activity$date.time <- as.POSIXct(paste(activity$date, time),
format='%Y-%m-%d %H.%M',
tz='GMT')
```
Analyzes the mean at different times of day, it is convenient to have a time column. So I convert all of the dates to be for today. since we only care about the time for that column, it will help with the analysis.
```{r}
activity$time <- format(activity$date.time, format='%H:%M:%S')
activity$time <- as.POSIXct(activity$time, format='%H:%M:%S')
```
## What is mean total number of steps taken per day?
Calculates the mean number of steps for each day:
```{r}
total.steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```
Mean and median for the total steps per day:
```{r}
mean(total.steps)
median(total.steps)
```
Distribution of total number of steps per day with a histogram:
```{r histogram}
library(ggplot2)
qplot(total.steps, xlab='Total steps', ylab='Frequency')
```
## What is the average daily activity pattern?
Calculates the mean steps for each five minute interval, and then put it in a data frame.
```{r}
mean.steps <- tapply(activity$steps, activity$time, mean, na.rm=TRUE)
daily.pattern <- data.frame(time=as.POSIXct(names(mean.steps)),
mean.steps=mean.steps)
```
Look at a time series plot for the mean steps.
```{r timeseriesplot}
library(scales)
ggplot(daily.pattern, aes(time, mean.steps)) +
geom_line() +
xlab('Time of day') +
ylab('Mean number of steps') +
scale_x_datetime(labels=date_format(format='%H:%M'))
```
Which five minute interval has the highest mean number of steps?
```{r}
most <- which.max(daily.pattern$mean.steps)
format(daily.pattern[most,'time'], format='%H:%M')
```
## Imputing missing values
Identify the number of intervals with missing step counts ("NA's"):
```{r}
summary(activity$steps)
```
To fill the missing values, use mean steps for a 5 minutes interval for the entire dataset.
```{r}
activity.imputed <- activity
activity.imputed$steps <- with(activity.imputed, activity.imputed$steps, mean)
```
Compares the mean and median steps for each day between the original data set and the imputed data set.
```{r}
total.steps.imputed <- tapply(activity.imputed$steps,
activity.imputed$date, sum)
mean(total.steps)
mean(total.steps.imputed)
median(total.steps)
median(total.steps.imputed)
```
And a histogram of the imputed dataset.
```{r histogram_imputed}
qplot(total.steps.imputed, xlab='Total steps', ylab='Frequency')
```
Input of the missing data has increased the average number of steps.
## Are there differences in activity patterns between weekdays and weekends?
Add a factor column for whether a day is a weekday or weekend.
```{r}
day.type <- function(date) {
if (weekdays(date) %in% c('Saturday', 'Sunday')) {
return('weekend')
} else {
return('weekday')
}
}
day.types <- sapply(activity.imputed$date.time, day.type)
activity.imputed$day.type <- as.factor(day.types)
```
Creates a dataframe that holds the mean steps for weekdays and weekends.
```{r}
mean.steps <- tapply(activity.imputed$steps,
interaction(activity.imputed$time,
activity.imputed$day.type),
mean, na.rm=TRUE)
day.type.pattern <- data.frame(time=as.POSIXct(names(mean.steps)),
mean.steps=mean.steps,
day.type=as.factor(c(rep('weekday'),
rep('weekend'))))
```
Compares patterns between weekdays and weekends.
```{r timeseries_daytype}
ggplot(day.type.pattern, aes(time, mean.steps)) +
geom_line() +
xlab('Time of day') +
ylab('Mean number of steps') +
scale_x_datetime(labels=date_format(format='%H:%M')) +
facet_grid(. ~ day.type)
```